Automating Verification of MPC Security

### 2023-06-20 popl24/bruteforce.tex
typos

Here we show how to rewrite any $\prog$ to an equivalent Datalog
program, which supports application of recent work in linear algebraic
interpretation of Datalog and optimizations of model computation on
high performance computers \cite{sakama2017linear}. The method here also enumerates
$\runs(\prog)$ memory-by-memory, rather than in a ``batched'' manner
as in our first method, allowing parallelization of model
computation. The rewriting we describe here is to Datalog with
negation, with a negation-as-failure model, though we can use also
use techniques in \cite{sakama2017linear} to eliminate negation from
resulting programs. \emph{Atoms} are $\minifed$ variables, \emph{literals}
are atoms or negated atoms, and clause bodies are conjunctions of literals.
A \emph{Datalog program} is a list of clauses.
$$
\begin{array}{rclr}
  \alpha &::=& \view{\cid}{w} \mid  \secret{\cid}{w} \mid \flip{\cid}{w} \mid \oracle{w} & \qquad \textit{(atoms)}\\
  \mathit{body} &::=&  \alpha \mid \neg \alpha \mid \alpha \wedge \mathit{body} \mid \neg\alpha \wedge \mathit{body} \mid \varnothing \\
  \mathit{clause} &::=& \alpha \gets \mathit{body}
\end{array}
$$

>>>>

Here we show how to rewrite any $\prog$ to an equivalent Datalog
program, which supports application of recent work in linear algebraic
interpretation of Datalog and optimizations of model computation on
high performance computers \cite{sakama2017linear}. The method here also enumerates
$\runs(\prog)$ memory-by-memory, rather than in a ``batched'' manner
as in our first method, allowing parallelization of model
computation. The rewriting we describe here is to Datalog with
negation, with a negation-as-failure model, though we can also
use techniques in \cite{sakama2017linear} to eliminate negation from
resulting programs. \emph{Atoms} are $\minifed$ variables, \emph{literals}
are atoms or negated atoms, and clause bodies are conjunctions of literals.
A \emph{Datalog program} is a list of clauses.
$$
\begin{array}{rclr}
  \alpha &::=& \view{\cid}{w} \mid  \secret{\cid}{w} \mid \flip{\cid}{w} \mid \oracle{w} & \qquad \textit{(atoms)}\\
  \mathit{body} &::=&  \alpha \mid \neg \alpha \mid \alpha \wedge \mathit{body} \mid \neg\alpha \wedge \mathit{body} \mid \varnothing \\
  \mathit{clause} &::=& \alpha \gets \mathit{body}
\end{array}
$$

### 2023-06-20 popl24/composition.tex
checkpoint
typos

\emph{Wire labels} are fundamental to YGC, and essentially represent
gate output values in an encrypted form. In our definition, wire
labels are representated by records $\ttt{\{ k = }\beta_1\ttt{; p =
}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
flips. Flips in each output wire label are owned by the garbler and
are unique per gate by definition of their identifying string, and the
representation of $0$ is the negation of $1$. For example, here is the
representation of 1 and 0 respectively in the output wire label for a
hypothetical gate 6:
\begin{mathpar}
  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}

@@

In our implementation, gates are wired together using gate
identifiers, which are strings $w$. Top-level functionality in Figures
\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
\begin{itemize}
\item \ttt{andgate}: This defines a subprotocol for the garbler
  to define a garbled gate $\ttt{gid}$ with input wires from gates
  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
  the rows in YGC fashion, them with client $1$ in
  views in a standard form. For example, the view for
  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
  We note that garbled gates of other binary operators can be obtained with
  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
  $\ttt{wvb}$.
\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
  defines the garbler's protocol for encrypting the circuit
  output from final gate $\ttt{gid}$, and the latter defines
  the evaluator's output decryption protocol.
\item \ttt{encode}: This defines the initial phase of the protocol,
  where the evaluator receives the wire value from their own
  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
  the wire value for their own secret $\sx{2}{sb}$ directly.
\end{itemize}
\begin{example}
  \label{example-andcircuit}
The following program uses our YGC library to define
a ciruit with a single and gate and input secrets $\ttt{s1}$ and
$\ttt{s2}$ from client's 1 and 2 respectively. 
\begin{verbatimtab}
  andgate(0,s1,s2);
  garbledecode(0);
  let secrets = encode(s1,s2) in
  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
\end{verbatimtab}
\end{example}
We have verified passive security of the $\fedprot$ protocol
generated by this and other small circuits using the
technique described in Lemma \ref{lemma-bruteforce-nimo}.
But large circuits with thousands of gates would be
intractable to verify with this method. In the next Section
we discuss compositional methods to address this issue.

@@

A basic intuition about YGC is that garbled gates considered
individually are ``nearly'' passive secure, modulo the feature that
outputs of individual logic gates in circuits are encryped. That is,
each gate enjoys an encrypted variant of$\NIMO$ such that the relevant
output is not in the clear but correlated flips sampled for the gate's
output wire label by the garbler, both key and pointer bits. We call
this \emph{Garbled $\NIMO$}, or $\gNIMO$. In the definition of
$\gNIMO$ we consider whether a given value $\itv$ correlates
positively or negatively to the flip chosen for a wire lable, which
can be either $1$ or $0$ uniformly- recall that
Definition \ref{def-progd} posits a variable $\itv$ to refer to the
expression $\be$ when considering the distribution
$\progd(\prog,\be)$.
\begin{definition}[Garbled $\NIMO$]
  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} =
  \be_{\ttt{p}}\}, g)$ with $\iov(\prog) = S \cup V$ iff for all $H$,
  $C$ with $|C| \le |H|$, for all $\store \in \mems(S_C \cup V_C)$,
  for both $w \in \{\ttt{p},\ttt{k}\}$, and for all $\stores$ with
  domain $X = S_C \cup V_C \cup \{ \itv, \flip{2}{\ttt{gate:}g.w} \}$
  such that:
  $$
  \stores = \{ \store' \mid \store \subset \store' \wedge 
  \store'(\itv) = \store'(\flip{2}{\ttt{gate:}g.w}) \} 
  $$
  or
  $$
  \stores = \{ \store' \mid \store \subset \store' \wedge 
  \store'(\itv) = \neg\store'(\flip{2}{\ttt{gate:}g.w}) \} 
  $$
  %where for any $\beta$:
  %$$\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \beta \}$$
  %or
  %$$
  %\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \neg\beta \}
  %$$
  we have:
  $$
  \margd{(\condd{\progd(\prog,\be_w)}{\stores_{X-{V_C}}})}{S_H} =
  \margd{(\condd{\progd(\prog,\be_w)}{\stores})}{S_H}
  $$
\end{definition}

@@

We aim to show that gates are probabilistically independent of each
other (i.e., are separated), and this independence supports
correctness-preserving composition. A critical property is that
separation is preserved by safe (in the sense of Theorem
\ref{theorem-metalang-safety}) substitution of programs, assuming
disjointness of flips used in programs separated programs.
\begin{lemma}
  \label{lemma-scope}
  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and $\prog_1$
  covers $X$, then 
  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
  $\prog_1'$ such that $\vars(\prog_1') \cap \vars(\prog_2) = X$ and
  $\prog_1';\prog_2$ is safe.
\end{lemma}
\begin{proof}
  Suppose on the contrary that there exists $\prog_1'$ that is not
  separable from $\prog_2$. By results in
  \cite{barthe2019probabilistic}, this implies that there
  exists $\store^n$ and $\store^n$ for $0 \le n \le 1$ such that
  $\store^1_X = \store^{1'}_X$ where:
  $$
  \config{\store^0}{\prog_1;\prog_2} \redxs \config{\store^1}{\prog_2} \redxs
  \config{\store^2}{\varnothing}
  $$
  and:
  $$
  \config{\store^{0'}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
  \config{\store^{2'}}{\varnothing}
  $$
  where $\store^2_{\vars(\prog_2)} \ne \store^{2'}_{\vars(\prog_2)}$, which is a
  contradiction.
%   This implies that there is some
%  $\store \in \mems(\vars(\prog_1';\prog_2))$ such that
%  $\store_{\vars{\prog_{1}'}} \in \runs{\prog_1'}$ and
%  \mems(\vars(\prog_1))$ with   
\end{proof}

>>>>

\emph{Wire labels} are fundamental to YGC, and essentially represent
gate output values in an encrypted form. In our definition, wire
labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
flips. Flips in each output wire label are owned by the garbler and
are unique per gate by definition of their identifying string, and the
representation of $0$ is the negation of $1$. For example, here is the
representation of 1 and 0 respectively in the output wire label for a
hypothetical gate 6:
\begin{mathpar}
  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}

@@

In our implementation, gates are wired together using gate
identifiers, which are strings $w$. Top-level functionality in Figures
\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
\begin{itemize}
\item \ttt{andgate}: This defines a subprotocol for the garbler
  to define a garbled gate $\ttt{gid}$ with input wires from gates
  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
  the rows in YGC fashion, them with client $1$ in
  views in a standard form. For example, the view for
  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
  We note that garbled gates of other binary operators can be obtained with
  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
  $\ttt{wvb}$.
\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
  defines the garbler's protocol for encrypting the circuit
  output from final gate $\ttt{gid}$, and the latter defines
  the evaluator's output decryption protocol.
\item \ttt{encode}: This defines the initial phase of the protocol,
  where the evaluator receives the wire value from their own
  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
  the wire value for their own secret $\sx{2}{sb}$ directly.
\end{itemize}
\begin{example}
  \label{example-andcircuit}
The following program uses our YGC library to define
a circuit with a single and gate and input secrets $\ttt{s1}$ and
$\ttt{s2}$ from client's 1 and 2 respectively. 
\begin{verbatimtab}
  andgate(0,s1,s2);
  garbledecode(0);
  let secrets = encode(s1,s2) in
  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
\end{verbatimtab}
\end{example}
We have verified passive security of the $\fedprot$ protocol
generated by this and other small circuits using the
technique described in Lemma \ref{lemma-bruteforce-nimo}.
But large circuits with thousands of gates would be
intractable to verify with this method. In the next Section
we discuss compositional methods to address this issue.

@@

A basic intuition about YGC is that garbled gates considered
individually are ``nearly'' passive secure, modulo the feature that
outputs of individual logic gates in circuits are encrypted. That is,
each gate enjoys an encrypted variant of $\NIMO$ such that the relevant
output is not in the clear but correlated flips sampled for the gate's
output wire label by the garbler, both key and pointer bits. We call
this \emph{Garbled $\NIMO$}, or $\gNIMO$. In the definition of
$\gNIMO$ we consider whether a given value $\itv$ correlates
positively or negatively to the flip chosen for a wire label, which
can be either $1$ or $0$ uniformly- recall that
Definition \ref{def-progd} posits a variable $\itv$ to refer to the
expression $\be$ when considering the distribution
$\progd(\prog,\be)$.
\begin{definition}[Garbled $\NIMO$]
  We write $\gNIMO(\prog,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} =
  \be_{\ttt{p}}\}, g)$ with $\iov(\prog) = S \cup V$ iff for all $H$,
  $C$ with $|C| \le |H|$, for all $\store \in \mems(S_C \cup V_C)$,
  for both $w \in \{\ttt{p},\ttt{k}\}$, and for all $\stores$ with
  domain $X = S_C \cup V_C \cup \{ \itv, \flip{2}{\ttt{gate:}g.w} \}$
  such that:
  $$
  \stores = \{ \store' \mid \store \subset \store' \wedge 
  \store'(\itv) = \store'(\flip{2}{\ttt{gate:}g.w}) \} 
  $$
  or
  $$
  \stores = \{ \store' \mid \store \subset \store' \wedge 
  \store'(\itv) = \neg\store'(\flip{2}{\ttt{gate:}g.w}) \} 
  $$
  %where for any $\beta$:
  %$$\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \beta \}$$
  %or
  %$$
  %\store^2 = \store\{ \flip{2}{\ttt{gate:}g.w} \mapsto \beta \}\{ \itv \mapsto \neg\beta \}
  %$$
  we have:
  $$
  \margd{(\condd{\progd(\prog,\be_w)}{\stores_{X-{V_C}}})}{S_H} =
  \margd{(\condd{\progd(\prog,\be_w)}{\stores})}{S_H}
  $$
\end{definition}

@@

We aim to show that gates are probabilistically independent of each
other (i.e., are separated), and this independence supports
correctness-preserving composition. A critical property is that
separation is preserved by safe (in the sense of Theorem
\ref{theorem-metalang-safety}) substitution of programs, assuming
disjointness of flips used in programs separated programs.
\begin{lemma}
  \label{lemma-scope}
  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and $\prog_1$
  covers $X$, then 
  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
  $\prog_1'$ such that $\vars(\prog_1') \cap \vars(\prog_2) = X$ and
  $\prog_1';\prog_2$ is safe.
\end{lemma}
\begin{proof}
  Suppose on the contrary that there exists $\prog_1'$ that is not
  separable from $\prog_2$. By results in
  \cite{barthe2019probabilistic}, this implies that there
  exists $\store^n$ and $\store^n$ for $0 \le n \le 1$ such that
  $\store^1_X = \store^{1'}_X$ where:
  $$
  \config{\store^0}{\prog_1;\prog_2} \redxs \config{\store^1}{\prog_2} \redxs
  \config{\store^2}{\varnothing}
  $$
  and:
  $$
  \config{\store^{0}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
  \config{\store^{2'}}{\varnothing}
  $$
  where $\store^2_{\vars(\prog_2)} \ne \store^{2'}_{\vars(\prog_2)}$, which is a
  contradiction.
%   This implies that there is some
%  $\store \in \mems(\vars(\prog_1';\prog_2))$ such that
%  $\store_{\vars{\prog_{1}'}} \in \runs{\prog_1'}$ and
%  \mems(\vars(\prog_1))$ with   
\end{proof}

### 2023-06-20 popl24/hyperprop.tex
checkpoint
typos

\section{Hyperproperties of MPC Passive Security}

@@

\subsection{Passive Security and the Real/Ideal Model} Here we
briefly summarize the formal definition of MPC security. For a
more detailed discussion the reader is referred to \cite{XXX}.
In this paper we will consider the \emph{passive} aka
semi-honest aka honest-but-curious variant. In this threat
model, protocol participants can be corrupted, but all clients
follow the rules of the protocol- as opposed to the
\emph{active} aka maliciouos variant where participants
are allowed to break these rules. In the passive model, the
vulnerability is through information leaked to the adversary
through the views of corrupted participants. While the passive
model also assumes that any clients can potentially be
corrupted, it assumes that no more than half will be-
i.e., it assumes an \emph{honest majority}.

>>>>

\section{Hyperproperties of MPC Passive Security}
\label{section-hyperprop}

@@

\subsection{Passive Security and the Real/Ideal Model} \dnote{would shortening the background info on how simulator model works be a bad idea?} Here we
briefly summarize the formal definition of MPC security. For a
more detailed discussion the reader is referred to \cite{XXX}.
In this paper we will consider the \emph{passive} aka
semi-honest aka honest-but-curious variant. In this threat
model, protocol participants can be corrupted, but all clients
follow the rules of the protocol- as opposed to the
\emph{active} aka malicious variant where participants
are allowed to break these rules. In the passive model, the
vulnerability is through information leaked to the adversary
through the views of corrupted participants. While the passive
model also assumes that any clients can potentially be
corrupted, it assumes that no more than half will be-
i.e., it assumes an \emph{honest majority}.

### 2023-06-20 popl24/intro.tex
typos

In this paper we make the following contributions.
\begin{itemize}
\item A new probabilistic programming language $\minifed$ for defining
  synchronous distributed protocols over the binary field. We consider example
  protocols including OT, communications encryption with one-time
  pads, and additive secret sharing (Section \ref{section-minicat}).
  \cnote{I'm hoping $\minifed$ hasn't been used already.}
\item A formulation of $\minifed$ program distributions, including
  distributions of secrets and views, supporting expression of security
  properties (Section \ref{section-pmf}). This intensional characterization
  is useful for expressing internal characteristics of protocols, and
  is needed for MPC due to the fact that obliviousness alone is not sufficient
  to capture characteristics of the real/ideal model, in particular the
  allowance of some information leakage through public outputs. 
\item A novel hyperproperty of program execution traces, called
  \emph{nonintererence modulo output ($\NIMO$)}, based on our
  formulation of program distributions, that implies passive security
  (Section \ref{section-hyperprop}).
\item A brute force mechanism for verifying passive security in
  $\minifed$ protocols, that is also amenable to HPC optimizations via
  translation of $\minifed$ protocols into Datalog (Section \ref{section-bruteforce}). 
\item A new metaprogramming language $\metaprot$ that dynamically
  generates $\minifed$ protocols. It includes control and data structures
  and is able to express logical protocol components, and enjoys a
  type safety result that guarantees that generated protocols
  are semantically well-defined (Section \ref{section-metalang}).
\item A formulation of Yao's Garbled Circuits, and a formulation of a
  compositional property of an extensible library of garbled gate
  components that is verifiable automatically. This property
  guarantees that any well-formed circuits using the library are
  passive secure (Section \ref{section-composition}).
\end{itemize}

>>>>

In this paper we make the following contributions.
\begin{itemize}
\item A new probabilistic programming language $\minifed$ for defining
  synchronous distributed protocols over the binary field. We consider example
  protocols including OT, communications encryption with one-time
  pads, and additive secret sharing (Section \ref{section-minicat}).
  \cnote{I'm hoping $\minifed$ hasn't been used already.}
\item A formulation of $\minifed$ program distributions, including
  distributions of secrets and views, supporting expression of security
  properties (Section \ref{section-pmf}). This intensional characterization
  is useful for expressing internal characteristics of protocols, and
  is needed for MPC due to the fact that obliviousness alone is not sufficient
  to capture characteristics of the real/ideal model, in particular the
  allowance of some information leakage through public outputs. 
\item A novel hyperproperty of program execution traces, called
  \emph{noninterference modulo output ($\NIMO$)}, based on our
  formulation of program distributions, that implies passive security
  (Section \ref{section-hyperprop}).
\item A brute force mechanism for verifying passive security in
  $\minifed$ protocols, that is also amenable to HPC optimizations via
  translation of $\minifed$ protocols into Datalog (Section \ref{section-bruteforce}). 
\item A new metaprogramming language $\metaprot$ that dynamically
  generates $\minifed$ protocols. It includes control and data structures
  and is able to express logical protocol components, and enjoys a
  type safety result that guarantees that generated protocols
  are semantically well-defined (Section \ref{section-metalang}).
\item A formulation of Yao's Garbled Circuits, and a formulation of a
  compositional property of an extensible library of garbled gate
  components that is verifiable automatically. This property
  guarantees that any well-formed circuits using the library are
  passive secure (Section \ref{section-composition}).
\end{itemize}

### 2023-06-20 popl24/metalang.tex
checkpoint
typos

The syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot-syntax}.  It includes a syntax of function
definitions and records, and values include client ids, identifier
strings, and boolean expressions. Expression forms allow dynamic
construction of boolean expression forms and view assignments. When
$\metaprot$ programs construct a $\fedprot$ assignment, a side effect
occurs whereby the assignment is added to the end of the $\fedprot$
program accumulated during evaluation.

Formally, we consider a complete metaprogram to include both a
codebase and a ``main'' program that uses the codebase. 
\begin{definition}
A \emph{codebase} $\codebase$ is a list of function 
declarations. We write $ \codebase(f) = x_1,\ldots,x_n,\ e$
iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
A \emph{metaprogram}, aka \emph{mataprotocol} is a pair of a 
codebase and expression $\codebase, e$. We may omit
$\codebase$ if it is clear from context.  
\end{definition}

@@

We define a small-step evaluation aka reduction relation $\redx$ in
Figure \ref{fig-metaprot-semantics}.  We write $\redxs$ to denote the
reflexive, transitive closure of $\redx$. Reduction is defined on
\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
where $\prog$ is the $\minifed$ program accumulated during evaluation.
In this definition we write $e[v/x]$ to denote the substitution of $v$
for free occurences of $x$ in $e$. The rules are mostly standard,
except why a concrete $\minifed$ assignment is encountered it is added
to the end of $\prog$.

@@

It is desirable to statically enforce safety of both $\metaprot$
programs and the safety of the $\fedprot$ programs they
generate. Although safety of the latter could be enforced
post-generation by a direct analysis, for large programs this can be
much more expensive and it is also better to not wast time on
resource-intensive compilation of programs with known errors
\cite{kreuter2012billion}. Some consequences of safety errors, for example accidental
reuse of one-time pads, can also undermine security.

@@

\inferrule[\TirName{SecretT}]
{\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}}
{\tjudge{\viewst}{\Gamma}{\secret{e_1}{e_2}}{\bet{e_1'}}}

>>>>

The syntax of $\metaprot$ is defined in Figure
\ref{fig-metaprot-syntax}.  It includes a syntax of function
definitions and records, and values include client ids, identifier
strings, and boolean expressions. Expression forms allow dynamic
construction of boolean expression forms and view assignments. When
$\metaprot$ programs construct an $\fedprot$ assignment, a side effect
occurs whereby the assignment is added to the end of the $\fedprot$
program accumulated during evaluation.

Formally, we consider a complete metaprogram to include both a
codebase and a ``main'' program that uses the codebase. 
\begin{definition}
A \emph{codebase} $\codebase$ is a list of function 
declarations. We write $ \codebase(f) = x_1,\ldots,x_n,\ e$
iff $f(x_1,\ldots,x_n) \{ e \} \in \codebase$.
A \emph{metaprogram}, aka \emph{mataprotocol} \dnote{meta?} is a pair of a 
codebase and expression $\codebase, e$. We may omit
$\codebase$ if it is clear from context.  
\end{definition}

@@

We define a small-step evaluation aka reduction relation $\redx$ in
Figure \ref{fig-metaprot-semantics}.  We write $\redxs$ to denote the
reflexive, transitive closure of $\redx$. Reduction is defined on
\emph{configurations} which are pairs of the form $\config{\prog}{e}$,
where $\prog$ is the $\minifed$ program accumulated during evaluation.
In this definition we write $e[v/x]$ to denote the substitution of $v$
for free occurences of $x$ in $e$. The rules are mostly standard,
except why \dnote{should this be a when?} a concrete $\minifed$ assignment is encountered it is added
to the end of $\prog$.

@@

It is desirable to statically enforce safety of both $\metaprot$
programs and the safety of the $\fedprot$ programs they
generate. Although safety of the latter could be enforced
post-generation by a direct analysis, for large programs this can be
much more expensive and it is also better to not waste time on
resource-intensive compilation of programs with known errors
\cite{kreuter2012billion}. Some consequences of safety errors, for example accidental
reuse of one-time pads, can also undermine security.

@@

\inferrule[\TirName{SecretT}]
{\tjudge{\viewst}{\Gamma}{e_1}{\cidty{e_1'}}{\viewst_1}\\
\tjudge{\viewst_1}{\Gamma}{e_2}{\stringty{e_2'}}{\viewst_2}}
{\tjudge{\viewst}{\Gamma}{\secret{e_1}{e_2}}{\bet{e_1'}}{\views_2}}
